Refined Flare Rates for Stars Younger than 250\,Myr

### 2023-09-22 src/tex/ms.tex
init commit v3

% Title
\title{An open source scientific article}

% Author list
\author{@afeinstein20}

% Abstract with filler text
\begin{abstract}
    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
    Phasellus eu tellus sit amet tortor gravida placerat.
    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
    Praesent eget sem vel leo ultrices bibendum.
    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
    Curabitur auctor semper nulla donec varius orci eget risus.
    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
    Duis eget orci sit amet orci dignissim rutrum.
\end{abstract}

@@

Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
Pellentesque cursus luctus mauris \citep{Luger2021}.

>>>>

% Title
\title{Refined Flare Rates for Stars Younger than 250\,Myr}

% Author list
\author{Adina~D.~ Feinstein}
\altaffiliation{NHFP Sagan Fellow}
\affiliation{Laboratory for Atmospheric and Space Physics, University of Colorado
Boulder, UCB 600, Boulder, CO 80309}

% Abstract with filler text
\begin{abstract}
    Photometric exoplanet detection missions have revolutionized our understanding
    of stellar activity. Stellar flares are short-duration high-energy radiation
    outbursts which can be readily seen across young (< 800\,Myr) stars and even
    in high counts on mature M dwarfs. Here, we present the results of measuring
    stellar flare rates for very young stars, with ages $4.5 - 200$\,Myr right
    after they formed to understand the early evolution of magnetic activity. We
    have selected a sample of 5,700 stars which belong to 26 associations,
    nearby young moving groups, or clusters. We identified $\sim 29,000$ flares
    using the \texttt{stella} machine-learning algorithm. We measured the flare-
    frequency distribution slopes as a function of age and stellar effective
    temperature.
\end{abstract}

@@

This paper is presented as follows.

\section{Methodology}
\label{sec:methods}

\subsection{Sample Selection}

We aim to measure the evolution in flare rates as a function of stellar age for
stars with $1 \leq t_\textrm{age} \leq 250$\,Myr. We used the MOCA Data Base
(Gagné et al. in prep.) to identify nearby young moving groups, associations,
and open clusters with known ages which satisfied our criteria. We identified
unique 26~groups from which we created our sample list. We considered all targets
that are either bona fide members, high-likelihood candidate members, or candidate members.
This resulted in a catalog of 30,889 stars across 26 associations. We summarize
the sample and quote adopted ages for each association
in Table~\ref{tab:sample}.

\begin{deluxetable}{l r r r r r}[!ht]
\tabletypesize{\footnotesize}
\tablecaption{Adopted Ages of each Young Stellar Population and Number of
Stars per Group \label{tab:sample}}
\tablehead{
\colhead{Population} & \colhead{Age [Myr]} & \colhead{N$_\textrm{stars}$} &
\colhead{Ref.}
}
\startdata
AB Doradus  &  133$_{-20}^{+15}$  &  81  & 1 \\
Blanco 1 & 137.1$^{+7.0}_{-33}$ & 301 & 8\\
Carina  &  45  & 91  & 2\\
Carina-Musca & 32 & 22 & 5\\
Chamaeleon  &  5  &  366  & 3 \\
Columba  &  42  &  126  & 2\\
Greater Taurus Subgroup 5 & 8.5 & 56 & 5\\
Greater Taurus Subgroup 8 & 4.5 & 114 & 5\\
Lower Centaurus Crux &  15  & 590  & 4\\
MELANGE-1 & 250$_{-70}^{+50}$ & 22 & 10\\
Octans  &  35$\pm 5$ &   98  & 6 \\
Pisces Eridanis  &  120  &   176  & 7 \\
Pleiades & 127.4$^{+6.3}_{-10}$ & 1304 & 8\\
$\alpha$ Persei  &  79$_{-2.3}^{+1.5}$  &  467  & 8 \\
IC 2602 corona &  52.5$_{-3.7}^{+2.2}$  & 10  & 8 \\
IC 2602 system &  52.5$_{-3.7}^{+2.2}$  & 138  & 8 \\
NGC 2451A & 48.5 & 44 & 5\\
Oh 59 & 162.2 & 50 & 9\\
Platais 9 & 50 & 99 & 11\\
RSG2 & 126 & 87 & 12\\
Theia 301  &  195  &  372  & 9\\
Theia 95  &  30.2  &  191  & 9\\
TW Hydrae  &  10  &  24  & 2\\
Upper Centaurus  Lupus &  16  &  464  & 4 \\
Upper Scorpius  &  10  &  80  & 4 \\
Vela-CG4  &  33.7  &  352  & 5
\enddata
\tablecomments{Age references: (1) \citet{gagne18_abdmg}; (2) \citet{Bell15};
(3) \citet{luhman07}; (4) \citet{pecaut16}; (5) \citet{kerr21};
(6) \citet{murphy15}; (7) \citet{curtis19}; (8) \citet{galindo22};
(9) \citet{kounkel20}; (10) \citet{tofflemeire21}; (11) \citet{tarricq21};
(12) \citet{roser16}; }
\end{deluxetable}

\subsection{TESS Light Curves}

To ensure we were able to resolve and accurately measure the energies of flares
on these stars \citep{howard22}, we used the 2-minute data from the TESS mission.
We crossmatched our sample with the TESS Input Catalog (TIC) based on their RA and Dec.
We considered a star to be in the catalog if the distance between the target and
the nearest TIC was within $1"$. This resulted in a final sample of 5,725 unique targets
which have been observed at 2-minute cadence between Sector 1 and Sector 67, the
latest available sector upon this analysis. Given the proximity of many targets towards the
northern and southern ecliptic poles (Figure~\ref{fig:sample}), many targets were observed
over multiple sectors. We downloaded all available data, yielding a total of 17,964
light curves.

\begin{figure}[ht!]
    \script{sample.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/sample.pdf}
        \caption{
            Distribution of our selected sample across the sky and colored by the
            adopted age of the association (see Table~\ref{tab:sample}). The all-sky
            coverage by TESS has unlocked new populations of stars to observe.
            We take advantage of this observing strategy to measure flare rates
            across 26 different nearby young moving groups, clusters, and associations.
        }
        \label{fig:sample}
    \end{centering}
\end{figure}

\subsection{Flare Identification}

We followed the machine learning flare-identification methods presented in
\cite{feinstein20a}. This method relies on the fact that all flare events
have similar time-dependent morphologies generally described as a sharp rise
and an exponential decay. This method uses a convolutional neural network (CNN),
\texttt{stella} \citep{feinstein20b}, trained on by-eye validated flare events from
TESS Sectors 1 and 2 to identify flare events in TESS 2-minute data
\citep{guenther19_flares, feinstein20a}. The benefits of the CNN is that it is
insensitive to the stellar baseline, since it is trained to look only for the
flare morphology. This means that the peaks of rotational modulation driven by
stellar heterogeneities, which is readily seen in the light curves of young stars,
are not accidentally identified as flares. In this way, we are able to
build a sample of flares which is unbiased towards low-amplitude/low-energy flares,
which are often not identified in traditional sigma-outlier identification
methods \citep{chang15}.

The \texttt{stella} CNN models take the light curve (time, flux, flux error)
as an input and returns an array with values of [0,1], which are treated as
the probability a data point is (1) or is not (0) part of a flare. We ran all
light curves through 10 \texttt{stella} models and averaged the output, taking
this as the final prediction per observation. The \texttt{stella} code uses the
predictions per data point to group together points, identifying those as single
flare events. It then calculates the probability of the entire flare event as
the average of the probabilities assigned to all the data points during the flare.

\begin{figure}[ht!]
    \script{flare_distribution.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/flare_distribution.pdf}
        \caption{
            High level summary of the demographics of flares in our sample. Top:
            The number of flares identified compared to the number of stars in
            each nearby young moving group, cluster, or association. A one-to-one
            relationship is expected. Middle: The distribution measured TESS energies
            and equivalent durations of flares in our sample, colored by the probability
            of the flare as identified with \texttt{steela}. Bottom: Same as the
            middle plot, except colored by the $T_\textrm{eff}$ [K] of the star.
            We limit our sample to stars with $T_\textrm{eff} \leq 6000$\,K.
        }
        \label{fig:flare_distribution}
    \end{centering}
\end{figure}

\subsection{Modeling Flare Properties}

We used the analytical flare model in
\cite{tovar22}\footnote{\url{https://github.com/lupitatovar/Llamaradas-Estelares}}
to fit the flares in our sample. This model builds upon the model presented
in \cite{davenport14}, and includes a convolution of a Gaussian with a
double exponential model. The analytical model accounts for the amplitude,
heating timescale, rapid cooling phase timescale, and slow cooling phase
timescale of flares. Using this model, we used a non-linear least squares
optimization to fit for the time of the flare peak ($t_\textrm{peak}$),
full width at half maximum (FWHM), and the amplitude ($A$), of each flare
in the sample. We combine the model with a second-order polynomial fit to
a 10-hour baseline before and after the flare to account for any rotational
modulation.

From these fits, we performed a series of quality checks to ensure the flares
in our sample are real. First, we removed any flares with best-fit FWHM $ < 0$
and $A < 0$. Second, we removed any flares with a large error on the amplitude,
$\sigma_A > 10$ and flares where $\sigma_A > A$. Third, we removed flares where
$A < 2 \times \textrm{RMS}$, where the RMS is calculated from the 10~hour baseline
around the best-fit $t_\textrm{peak}$. These cuts are indicative of a poorly fit
flare model due to the flare being overwhelmed by noise in the surrounding baseline
light curve. We choose to include flares with a probability $P \geq 75\%$ of being
a true flare. After these checks, we have a flare sample of 29,128~flares originating
from 3,983 stars (Figure~\ref{fig:flare_distribution}). 83.6\% of the flares have
probabilities of $P \geq 90\%$ of being true; 64.9\% of the flares have probabilities
of $P \geq 99\%$ of being true.

\section{Results}\label{sec:res}

\begin{figure*}[ht!]
    \script{mcmc_results.py}
    \begin{centering}
        \includegraphics[width=2\textwidth]{figures/mcmc_results.pdf}
        \caption{
            Measured flare-frequency distribution slopes, $\alpha$, as a function
            of stellar effective temperature, $T_\textrm{eff}$ and age.
        }
        \label{fig:mcmc_results}
    \end{centering}
\end{figure*}

\subsection{Flare Rates by Association}

\subsection{Flare Rates by Age}

\subsection{Flare Rates by Effective Temperature}

\section{Conclusions}\label{sec:conclusions}
